Research on The Impact of Past Behaviour Normality on Regret

### 2024-02-14 outputs/paper/paper.qmd
introduction modified

Descriptive norms and injunctive norms are key concepts in understanding social behavior, each describing different aspects of how individuals perceive and are influenced by the norms within their community or society. Descriptive norms are about observing and following what most people do in a given situation, essentially reflecting the common actions or behaviors within a group. Injunctive norms, on the other hand, are about what people believe they should or shouldn't do, based on the group's values or expectations for behavior. These norms carry a sense of approval or disapproval, guiding behavior by suggesting what is socially acceptable or frowned upon. Both types of norms are powerful in shaping individual actions by providing a roadmap of 'what is done' and 'what ought to be done,' helping maintain social order and cohesion by aligning personal conduct with group expectations.

This study delves into how people make decisions, feel about their choices, and, how they deal with regret when things don't go as planned. How much more would people regret their decisions just because they're not what you usually do? This idea stems from a concept called norm theory, which suggests that people feel extra bad about outcomes when they result from actions that are not typical for them. Two main experiments was introduced to unravel this complex bundle of decisions and emotions. The first experiment places people in a scenario where they decide whether or not to pick up a hitchhiker, testing how much one would regret the decision based on whether it's something they normally do. The second experiment changes the setting to driving, where choosing a different route leads to an accident, and looks at how this break from routine affects their feelings of regret.

By picking apart these scenarios, this paper aims to shed light on the tricky relationship between stepping out of one's comfort zone in decision-making and the regret that can follow, especially when things don't turn out well. The goal of this research is to contribute to the broader conversation on how people's minds work in these situations, improving understanding of why people feel the way they do after making certain choices. This insight is not just for psychologists but for anyone interested in the little decisions people make every day and how they shape their feelings and lives. With all been said, we predicted that someone who often picks up hitchhikers would be seen as doing something unusual (against descriptive norms) and might be judged more harshly by others (against injunctive norms). We also introduced a new aspect to the original experiment, looking at how these social norms and a person's usual actions stir up feelings, both good and bad.

The remainder of this paper is structured as follows.

@sec-data
@sec-res
@sec-dis

>>>>

Regret, defined as an emotional response to the memory of past decisions or actions that one wishes had been different, significantly impacts our daily lives. This paper delves into the factors that influence the intensity of regret, with a specific focus on social norms and past behaviors. Understanding these elements is essential for a clearer grasp of the decision-making process and the emotions that result from it.

We categorize social norms into two main types: descriptive norms, which concern beliefs about common behaviors within a group, and injunctive norms, which relate to beliefs about what behaviors are approved or disapproved by others. Our research involves two experiments aimed at assessing the effect of these norms on regret. The first experiment looks into the role of these norms in the context of deciding whether or not to pick up a hitchhiker, while the second addresses the consequences of selecting an unconventional driving route.

The findings from these experiments reveal that unusual or unexpected events tend to provoke stronger feelings of regret than routine behaviors. However, the influence of both injunctive and descriptive social norms on the level of regret was found to be less significant. These results highlight how deviations from usual behavior, particularly when in conflict with social norms, can intensify feelings of regret. This insight is crucial for a better understanding of the cognitive processes involved in evaluating decisions and the emotional aftermath.

The organization of this paper is methodical and clear. We start with an overview of the theoretical framework concerning social norms and their connection to regret in the introduction. The methodology and findings from our experiments are detailed in the subsequent data section. The discussion section then addresses the implications of our results, outlines the limitations of the study, and suggests directions for future research.
